---
title: "Practice Diet Full"
author: "Roel"
date: "6/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pander)
options(digits=3)
```

# Practice Diet Full

To investigate the effect of 3 types of diet, 78 persons were divided randomly in 3 groups, the
first group following diet 1, second group diet 2 and the third group diet 3. Next to some other
characteristics, the weight was measured before diet and after 6 weeks of diet for each person
in the study. The collected data is summarized in the data frame diet.txt with the following
columns: person -- participant number, gender -- gender (1 = male, 0 = female), age -- age (years),
height -- height (cm), preweight -- weight before the diet (kg), diet -- the type of diet followed,
weight6weeks -- weight after 6 weeks of diet (kg).

```{r}
diet=read.table("diet.txt",header=TRUE);

```

## a) By using only the columns preweight and weight6weeks, test the claim that the diet affects
the weight loss. Give and check the assumptions of the test(s) applied.

```{r}

## a) By using only the columns preweight and weight6weeks, 
##test the claim that the diet affects
##the weight loss. Give and check the assumptions of the test(s) applied.

# x = preweight, y = postweight
x=diet$preweight;y=diet$weight6weeks

# we intend to use paired t-test, first check normality
# make space for 2 plots
par(mfrow=c(1,2))

#check normality assumption
qqnorm(x);hist(x) # does not look normal
qqnorm(y);hist(y) # does not look normal


qqnorm(y-x);hist(y-x) #shapiro.test(y-x) # but the difference is actually ok
t.test(x,y,alt="g",paired=T) # test says diet works, p < 0.05

## The samples x, y do not look normal, we can also use Mann-Whitney test

wilcox.test(x,y, alt = 'g', paired=T) # also this test says diet works 
# one can set in wilcoxon test alt="g", but it's fine without
```

## b) Compute and add to the data frame the variable weight.lost expressing the lost weight.
Test the claim that the median lost weight is bigger than 3 kg.

```{## b) Compute and add to the data frame the variable weight.lost}
## expressing the lost weight.
## Test the claim that the median lost weight is bigger than 3 kg.


# compute and add weight.lost
diet$weight.lost=diet$preweight-diet$weight6weeks

# We reduce the testing problem to a binomial test, 
# let X be the lost weight.  

# Notice that...
# H0: med(X)<= 3 
# H1: med(X)>3

# ... is the same as... 

# H0: p=P(X>3)<=0.5
# H1: p>0.5  


# The latter is a binomal test. So, 

#n = observed values
#w = sum of all >3 values
n=length(diet$weight.lost); w=sum(diet$weight.lost>3)


binom.test(w,n,0.5,alt="g") # p-value=0.04439<0.05, 
#we reject H0, hence the claim med(X)>3 is correct.
```

## c) Using now weight.lost as response variable, apply one-way ANOVA to test whether type
of diet has an effect on the lost weight. Do all three types diets lead to weight loss? Which
diet was the best for losing weight?

```{r}
## c) Using now weight.lost as response variable, 
## apply one-way ANOVA to test whether type of diet 
## has an effect on the lost weight. 
## Do all three types diets lead to weight loss? 
## Which diet was the best for losing weight?

diet$diet=factor(diet$diet) # make the variable diet factor
## is.factor(diet$diet) # just to check that it became factor
mod1=lm(weight.lost~diet,data=diet)
anova(mod1) ## p=0.003229, diet has an effect on the lost wight

par(mfrow=c(1,2)) # check the normality assumption
qqnorm(residuals(mod1)) # looks ok
plot(fitted(mod1),residuals(mod1)) # looks ok, remember there are just 3 fitted values

## Do all three types of diets lead to weight loss? 
# Yes, as all cell means mu+alpha_i>0 are positive.

# Which diet was best for losing weight?   
summary(mod1) # diet 3 is the best

```

## D) Use two-way ANOVA to investigate effect of the diet and gender (and possible interaction)
on the lost weight. Is the Friedman test relevant here? Can it be at all applied?

```{r}
## d)
# Test for interaction between gender and diet
mod2=lm(weight.lost~gender*diet,data=diet)  
anova(mod2) # interaction between gender and diet is barely present 
## it seems that diet is certainly significant, 
# gender is present as well at least via interaction which is there  

# Why can we not apply the Friedman test? 
# One simple (but not full) answer is: 
# the Friedman test is not relevant 
# because there is no block factor involved here. 
# We can pretend that gender is a block, 
# the test still cannot be applied in view of 
# the wrong design

# Additional reasoning why the Friedman test is not relevant. If we try
friedman.test(diet$weight.lost,diet$diet,diet$gender) 
# we see the problems with NA's
diet2=diet[-c(25,26),] # but even if we remove the rows with NA's, there will 
# still be a problem with the wrong design: we do not have one observations 
# for all combinations of levels of the treatment and block factors
friedman.test(diet2$weight.lost,diet2$diet,diet2$gender) # wrong design


```

## E) Apply appropriate model to investigate effect(s) of diet and height (and possibly their
interaction) on the lost weight. Is the effect of height the same for all 3 types of diet?

```{r}
## e)
is.numeric(diet$height) # just to check height is numeric
mod4=lm(weight.lost~diet*height,data=diet)
anova(mod4) # no interaction between diet and height

## Is the effect of height the same for all 3 types of diet?
# Yes, because there is no interaction between diet and height. 
# Now we test for the main effects by using the additive model 
anova(lm(weight.lost~diet+height,data=diet))
# conclude that diet is significant and height is not. 

```

## 
F) Which of the two approaches, the one from c) or from e), do you prefer? Why? For the
preferred model, predict the lost weight for all three types of diet for an average person.

```{r}
## f) Which of the two approaches, c) or e), do you prefer? Why? 
# For the preferred model, 
# predict the lost weight for all three types of diet for an average person.


# In e) we found there is no interaction. So we use the additive model
mod5=lm(weight.lost~diet+height,data=diet)
drop1(mod5,test="F") # factor diet is significant, variable height is not

## not asked: also summary(mod5) 
# confirms what we had in c)
# Hence the approach from c) is preferable (with only factor diet): 
# no reason to include variable height that has no effect whatsoever. 

# Prediction of the lost weight for all 3 types of diet for an average person:
# the characteristics of a person do not matter as they are not included in our preferred model (just one-way anova with factor diet)

summary(mod1) # and compute the group means: 3.3000kg, 3.3000-0.2741=3.0259kg, 
# 3.3000+1.8481=5.1481kg are expected to loose as result of diets 1,2,3
# Alternatively, by predict command:
newdata=data.frame(diet=c("1","2","3"))
predict(mod1,newdata) # 3.300000kg 3.025926kg 5.148148kg 

```

## G) Create the categorical (binary) variable lost.4kg expressing whether the lost weight is bigger
than 4kg or not. (Hint: if a=c(1,-1,2), then a\>0 = TRUE FALSE TRUE.)
Suppose now that we observed the variable lost.4kg instead of weight6weeks. Could we
still test the hypothesis in b), c), d), e)? If yes, propose relevant models and perform relevant
tests.

```{r}

## g) 

# Test in b) is impossible to implement

# Test in c) for lost.4kg and diet can be done in two ways:  
# either by contingency table test; or by using logistic regression.
# Any of the two is ok. We use logistic regression for the test in c).

# Test in c) by logistic regression lost.4kg~diet
lr1=glm(lost.4kg~diet,family=binomial,data=diet)
anova(lr1,test="Chisq") # diet is significant

# Tests in d) and e) can be addressed by the logistic regression
## Test in d) by logistic regression lost.4kg~diet*gender
lr3=glm(lost.4kg~gender*diet,family=binomial,data=diet)
anova(lr3,test="Chisq") # there is an interaction between diet and gender

# Test in e) by logistic regression lost.4kg~diet*height
lr5=glm(lost.4kg~diet*height,family=binomial,data=diet)
anova(lr5,test="Chisq") # no interaction between diet and height
lr6=glm(lost.4kg~diet+height,family=binomial,data=diet) # additive model
drop1(lr6,test="Chisq") # diet is significant, height is not
# or: anova(lr1,lr6,test="Chisq")

```
